---
title: "reproducible_research_project1"
author: "KBP"
date: "16 czerwca 2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Data loading
Load the data usinf read.csv()

```{r load}
activity <- read.csv("activity.csv")
```

## 2. Preprocess data
First prepare dates value to be real date format (makes plotting easier)

```{r data_prep}
activity$date <- as.Date(as.character(activity$date), "%Y-%m-%d")
```

## 3. Ignore missing data
### 3.1 Calculate steps taken each day
```{r calculate_total_steps}
sum_steps <- function(x){sum(x$steps, na.rm=TRUE)}
steps_each_day <- sapply(split(activity, activity$date),sum_steps)
steps_each_day
```

### 3.2 Plot histogram of of steps taken each day
```{r hist_of_steps}
hist(steps_each_day,main="Histogram of the total number of steps taken each day", xlab = "Number of steps taken per day")
```

### 3.3 Calculate mean of total steps taken per day
```{r mean_of_steps}
mean(steps_each_day)
```

### 3.4 Calculate median of total steps taken per day
```{r median_of_steps}
median(steps_each_day)
```

### 3.5 Average daily activity pattern - time series plot
```{r time-series_plot}
my_activity <- split(activity, activity$interval)
my_means <- matrix(nrow=length(my_activity),ncol=2)
for (i in 1:length(my_activity)){my_means[i,2]<-mean(my_activity[[i]]$steps, na.rm=TRUE)
my_means[i,1]<-unique(my_activity[[i]]$interval)}
plot(my_means, type="l", main="Averaged number of steps taken per interval", xlab="Interval", ylab="Number of steps")

```

### 3.6 Average daily activity pattern - interval with max number of steps
```{r averag_max_numb_steps}
my_means[which.max(my_means[,2]),1]
```


## 4. Impute missing data 

### 4.1 Calculate total number of NAs in steps column
```{r sum_nas}
sum(is.na(activity$steps))
```

### 4.2 Get rid of NAs in steps variable by switching it with mean of all the available values for this specific interval: 

```{r impute_NAs}
activity_fixed=activity
for (j in 1:length(activity_fixed$steps)){if(is.na(activity_fixed$steps[j])) activity_fixed$steps[j]=my_means[which(my_means[,1]==activity_fixed$interval[j]),2]}
head(activity_fixed)
head(activity)

```

### 4.3 Plot histogram of steps taken each day

```{r hist of steps per day}
steps_each_day_fixed <- sapply(split(activity_fixed, activity_fixed$date),sum_steps)
hist(steps_each_day_fixed,main="Histogram of the total number of steps taken each day \n for data with NA imputed", xlab = "Number of steps taken per day")
```


### 4.4 Mean and median number of steps taken each day
```{r mean_and_median_of_steps}
mean(steps_each_day_fixed)
median(steps_each_day_fixed)
```

### 4.5 Differences in activity patterns between weekdays and weekends
Check whether day is a weekday or weekend (for Polish week day names)
```{r weekdays_weekends}
weekday_or_weekend <- function(x){if((weekdays(x)=="sobota") | (weekdays(x)=="niedziela")) "weekend" else "weekday"}
activity_fixed$day_type <- factor(sapply(activity_fixed$date, weekday_or_weekend))
activity_dayType <- split(activity_fixed, activity_fixed$day_type)
my_activity_weekday <- split(activity_dayType[["weekday"]], activity_dayType[["weekday"]]$interval)
my_activity_weekend <- split(activity_dayType[["weekend"]], activity_dayType[["weekend"]]$interval)
my_means_weekday <- matrix(nrow=length(my_activity_weekday),ncol=2)
for (i in 1:length(my_activity_weekday)){my_means_weekday[i,2]<-mean(my_activity_weekday[[i]]$steps, na.rm=TRUE)
my_means_weekday[i,1]<-unique(my_activity_weekday[[i]]$interval)}
my_means_weekend <- matrix(nrow=length(my_activity_weekend),ncol=2)
for (i in 1:length(my_activity_weekend)){my_means_weekend[i,2]<-mean(my_activity_weekend[[i]]$steps, na.rm=TRUE)
my_means_weekend[i,1]<-unique(my_activity_weekend[[i]]$interval)}
 split.screen(c(2, 1))
 screen(1)
plot(my_means_weekday, type="l", main="Weekdays", xlab="",ylab="Number of steps")
screen(2)
plot(my_means_weekend, type="l", main="Weekend", xlab="Interval", ylab="Number of steps")
close.screen(all = TRUE)

```